%  Voorwoord (dankwoord) en toelating tot bruikleen

\begin{slshape}
\subsection*{Foreword}
\renewcommand{\baselinestretch}{1.2}
\small\normalsize

\npar 
After a long and sporadically lonesome road, an incredible journey has come to an end in the form of this master dissertation. Exploiting experience from exploration is what makes one fail less and less. \\

I had the amazing opportunity to pursue my thesis in a subject I am extremely passionate about, for which I want to thank my supervisor Prof. Francis wyffels.\\

Last year and especially the last three months, I needed a lot of advice and help. Luckily, Jeroen Burms has always been a great counselor and helped me whenever I had a question or idea. Thank you.\\

I want to express my gratitude to 'grand papa et grand maman'. During the fortnight of my isolation, they spoiled me more than my facebook wall spoiled the Game of Thrones story.\\

Furthermore I want to show appreciation to my fellow students with whom I spent countless hours of hard labor. I am thanking Jonas for lifting me up, Loic for always remaining casual, Timon for always being equally happy and Vincent for infecting invincibility.\\

Closer to me, there are my parents whom I love very much. I am sincerely sorry if I do not show this enough.\\

\noindent Dorian Van den Heede, august 2017
\end{slshape}

\newpage
\section*{Permission of use on loan}

\vspace{1.5cm}

\noindent
``The author(s) gives (give) permission to make this master dissertation available for consultation
and to copy parts of this master dissertation for personal use.
 In the case of any other use, the copyright terms have to be respected, in particular with regard to
the obligation to state expressly the source when quoting results from this master dissertation.''

\addvspace{4cm}

\noindent Dorian Van den Heede, august 2017
